---
title: "Synthetic Control-CA Tobacco"
author: "YC"
date: "2024-06-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synthetic Control tutorial

This tutorial aims to reproduce the results of Abadie et al. 2010, which evaluates the impact of Proposition 99 on cigarette consumption in California. This tutorial will use the `Synth` package in R, which implements synthetic control methods used in comparative case studies. These methods construct a synthetic control unit as weighted average of available control units that best approximate the relevant characteristics of the treated unit prior to the treatment.

```{r}
library(tidyverse)
library(skimr)
library(Synth)
library(SCtools)
library(rio)
```

## Data

Read in the [smoking example dataset](https://github.com/edunford/tidysynth/tree/master/data)

```{r}
dt_smoke = import("./data/smoking.rda")
```
### Data exploration

```{r}
glimpse(dt_smoke)
```

```{r}
skim(dt_smoke)
```

### Prep the data

```{r}


dt_smoke$st = as.numeric(as.factor(dt_smoke$state))
ctr_doners = unique(dt_smoke$st)
ctr_doners = ctr_doners[ctr_doners != 3]


smoke_out = dataprep(
  foo = dt_smoke,  # dataset to be prepared
  
  predictors = c("lnincome", "retprice", "age15to24"),
  predictors.op = "mean", # operation on the predictors
  time.predictors.prior = 1980:1988,
  
  # Special predictors with repective time periods and operations
  special.predictors = list(
    list("beer", 1984:1988, "mean"),
    list("cigsale", 1975, "mean"),
    list("cigsale", 1980, "mean"),
    list("cigsale", 1988, "mean")
  ),
  
  dependent = "cigsale",
  unit.variable = "st",
  unit.names.variable = "state",
  
  time.variable = "year",
  treatment.identifier = 3,
  controls.identifier = ctr_doners,
  
  time.optimize.ssr = 1970:1988,
  time.plot = 1970:2000
  
)


```

```{r}
# Treatment unit predictors
smoke_out$X1
```

```{r}
# Outcome variables BEFORE treatment
smoke_out$Z1
```
## Running the model

To get the optimal weights

```{r}
 synth_out = synth(data.prep.obj = smoke_out,
                    method = "BFGS" #   set optim() to use the BFGS quasi-Newton algorithm. 
                    )

```
## Results

Compares pre-treatment predictor values for the treated unit, the synthetic control unit, and all the units in the sample.

```{r}

synth_tables = synth.tab(dataprep.res = smoke_out,
                           synth.res    = synth_out
                           )
synth_tables$tab.pred
```
The weight corresponding to each predictor

```{r}
synth_tables$tab.v
```
The weight corresponding to each potential control unit

```{r}
synth_tables$tab.w
```

### Plot

Treatment vs synthetic control

```{r}
path.plot(synth.res = synth_out,
           dataprep.res = smoke_out,
           Ylab = "Cigar sales",
           Xlab = "year",
           Legend = c("California","synthetic control"),
           Legend.position = "bottomright"
           )
```

Gaps plot

```{r}

gaps.plot(synth.res = synth_out,
           dataprep.res = smoke_out,
           Ylab = "Cigar sales",
           Xlab = "year"
           )

```


### References
Abadie, A., Diamond, A., & Hainmueller, J. (2010). Synthetic Control Methods for Comparative Case Studies: Estimating the Effect of California’s Tobacco Control Program. Journal of the American Statistical Association, 105(490), 493–505. https://doi.org/10.1198/jasa.2009.ap08746

